---
title: "Optimal intervention model"
author: "Jacqueline Harding, Tobias Gerstenberg & Lara Kirfel"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: cosmo
    highlight: tango
---

# Load packages 

```{r, message=F}
library("knitr")
library("janitor")
library("patchwork")
library("Metrics")
library("tidyverse")
```

# Set options 

```{r}
theme_set(theme_classic() + 
    theme(text = element_text(size = 24)))

opts_chunk$set(comment = "",
               fig.show = "hold")

# suppress grouping warning 
options(dplyr.summarise.inform = F)
```

# Functions

```{r}
fun.softmax = function(x, temp = 3) {
    out = exp(x*temp) / sum(exp(x*temp))
    return(out)
}
```

# DATA 

```{r, warning=FALSE, message=FALSE}
df.data = bind_rows(
  read_csv(file = "../../../data/aggregate/hardint_pos.csv"),
  read_csv(file = "../../../data/aggregate/hardint_neg.csv"),
  read_csv(file = "../../../data/aggregate/softint_pos.csv"),
  read_csv(file = "../../../data/aggregate/softint_neg.csv"),
  read_csv(file = "../../../data/aggregate/fixedint_pos.csv"),
  read_csv(file = "../../../data/aggregate/fixedint_neg.csv")) %>% 
  clean_names() %>% 
  mutate(causal_structure = str_to_lower(causal_structure),
         experiment = str_remove(experiment, "int")) %>% 
  rename_with(.fn = ~str_remove_all(., "_percentage")) %>% 
  pivot_longer(cols = -c(causal_structure, outcome, experiment),
               values_to = "probability") %>% 
  separate(col = name,
           into = c("choice", "type")) %>% 
  mutate(across(.cols = -probability,
                .fns = ~ as.factor(.)),
         choice = factor(choice, levels = c("abnormal", "nopreference", "normal"))) %>%
  mutate(probability = probability / 100)

df.intervention = df.data %>% 
  filter(type == "intervention") %>% 
  rename(intervention = experiment)

df.explanation = df.data %>% 
  filter(type == "explanation")

colnames(df.data)

```



# MODEL 

## Interventions 

### Model structure

```{r}

fun.success = function(p_abnormal, p_normal, causal_structure, outcome){ 
  if (causal_structure == "conjunctive"){
    p = p_abnormal *  p_normal
  } else{
    p = 1 - (1 - p_abnormal) * (1 - p_normal)
  }
  if (outcome == "negative"){
    p = 1 - p
  }
  return(p)
}

causal_structure = c("conjunctive", "disjunctive")
outcome = c("positive", "negative") 

df.model = expand_grid(causal_structure, outcome) %>% 
  mutate(p_abnormal = 0.2,
         p_normal = 0.8,
         int_hard_abnormal = ifelse(outcome == "positive", 1, 0),
         int_hard_normal = ifelse(outcome == "positive", 1, 0),
         int_soft_abnormal = ifelse(outcome == "positive",
                                  p_abnormal + 0.2,
                                  p_abnormal - 0.2),
         int_soft_normal = ifelse(outcome == "positive",
                                  p_normal + 0.2,
                                  p_normal - 0.2),
         int_fixed_abnormal = ifelse(outcome == "positive",
                                  0.9,
                                  0.1),
         int_fixed_normal = ifelse(outcome == "positive",
                                  0.9,
                                  0.1),
         p_success = pmap_dbl(.l = list(p_abnormal, 
                                        p_normal, 
                                        causal_structure, 
                                        outcome),
                              .f = ~ fun.success(..1, ..2, ..3, ..4)),
         p_success_int_hard_abnormal = pmap_dbl(.l = list(int_hard_abnormal, 
                                        p_normal, 
                                        causal_structure, 
                                        outcome),
                              .f = ~ fun.success(..1, ..2, ..3, ..4)),
         p_success_int_hard_normal = pmap_dbl(.l = list(p_abnormal, 
                                        int_hard_normal, 
                                        causal_structure, 
                                        outcome),
                              .f = ~ fun.success(..1, ..2, ..3, ..4)),
         p_success_int_soft_abnormal = pmap_dbl(.l = list(int_soft_abnormal, 
                                        p_normal, 
                                        causal_structure, 
                                        outcome),
                              .f = ~ fun.success(..1, ..2, ..3, ..4)),
         p_success_int_soft_normal = pmap_dbl(.l = list(p_abnormal, 
                                        int_soft_normal, 
                                        causal_structure, 
                                        outcome),
                              .f = ~ fun.success(..1, ..2, ..3, ..4)),
         p_success_int_fixed_abnormal = pmap_dbl(.l = list(int_fixed_abnormal, 
                                        p_normal, 
                                        causal_structure, 
                                        outcome),
                              .f = ~ fun.success(..1, ..2, ..3, ..4)),
         p_success_int_fixed_normal = pmap_dbl(.l = list(p_abnormal, 
                                        int_fixed_normal, 
                                        causal_structure, 
                                        outcome),
                              .f = ~ fun.success(..1, ..2, ..3, ..4)))
```

### Choice probabilities 

```{r}
# gives expected value for each intervention
df.choice = df.model %>% 
  select(causal_structure, outcome,
         contains("p_success_int")) %>% 
  pivot_longer(-c(causal_structure, outcome)) %>% 
  mutate(name = str_remove(name, "p_success_int_")) %>% 
  separate(name, into = c("intervention", "target")) %>% 
  pivot_wider(names_from = target,
              values_from = value) %>% 
  arrange(intervention, causal_structure) %>% 
  relocate(intervention) %>% 
  mutate(nopreference = 0.5 * abnormal + 0.5 * normal) %>% 
  pivot_longer(c(abnormal, normal, nopreference),
               names_to = "choice") %>% 
  mutate(choice = factor(choice, levels = c("abnormal", "nopreference", "normal")),
         across(.cols = c(intervention, causal_structure, outcome),
                .fns = ~ as.factor(.)))
```

### Fit softmax parameter

```{r}
fun.fit_temperature = function(df_data, df_prediction, temperature){
  df_prediction %>% 
    group_by(intervention, causal_structure, outcome) %>% 
    mutate(prediction = fun.softmax(value, temp = temperature)) %>% 
    ungroup() %>% 
    left_join(df_data,
              by = c("intervention", "causal_structure", "outcome", "choice")) %>% 
    summarize(loss = sum((prediction - probability) ^ 2)) %>% 
    pull(loss)
}

fit.temperature = optim(par = 10, 
                        fn = fun.fit_temperature,
                        method = "L-BFGS-B",
                        lower = 0, 
                        upper = 100,
                        df_data = df.intervention,
                        df_prediction = df.choice)

print(fit.temperature$par)
```
### Predictions

```{r}
df.prediction_intervention = df.choice %>% 
  group_by(intervention, causal_structure, outcome) %>%
  mutate(prediction = fun.softmax(value, temp = fit.temperature$par)) %>%
    left_join(df.data %>% 
                filter(type == "intervention") %>% 
                select(-type) %>% 
                rename(intervention = experiment),
              by = c("causal_structure", "outcome", "intervention", "choice"))
```

## Explanations 

### Model structure 

```{r}
df.prediction_explanation =  df.choice %>% 
  group_by(intervention, causal_structure, outcome) %>%
  mutate(truth = ifelse(choice == "nopreference", 1, 0)) %>%
    left_join(df.data %>% 
                filter(type == "explanation") %>% 
                select(-type) %>% 
                rename(intervention = experiment),
              by = c("causal_structure", "outcome", "intervention", "choice")) %>% 
  ungroup()
```

### Model fitting functions 

#### Combined model

```{r}
fun.fit_params = function(params, df_prediction){
  
  weight <- params[1]
  temperature <- params[2]
  
  df_prediction %>% 
    group_by(intervention, causal_structure, outcome) %>% 
    mutate(prediction = fun.softmax(weight * value + (1 - weight) * truth, temp = temperature)) %>% 
    ungroup() %>% 
    summarize(loss = sum((prediction - probability) ^ 2)) %>% 
    pull(loss)
}
```

#### Intervention only model 

```{r}
fun.fit_params_intervention_only = function(params, df_prediction){
  
  temperature <- params[1]
  
  df_prediction %>% 
    group_by(intervention, causal_structure, outcome) %>% 
    mutate(prediction = fun.softmax(value, temp = temperature)) %>% 
    ungroup() %>% 
    summarize(loss = sum((prediction - probability) ^ 2)) %>% 
    pull(loss)
}
```

#### Truth only model 

```{r}
fun.fit_params_truth_only = function(params, df_prediction){
  
  temperature <- params[1]
  
  df_prediction %>% 
    group_by(intervention, causal_structure, outcome) %>% 
    mutate(prediction = fun.softmax(truth, temp = temperature)) %>% 
    ungroup() %>% 
    summarize(loss = sum((prediction - probability) ^ 2)) %>% 
    pull(loss)
}
```

### Fit paramters 

#### Combined model

```{r}

initial_params <- c(weight = 0.5, temperature = 10)
lower_bounds <- c(weight = 0, temperature = 0)
upper_bounds <- c(weight = 1, temperature = 100)

fit.params <- optim(par = initial_params, 
                         fn = fun.fit_params,
                         method = "L-BFGS-B",
                         lower = lower_bounds, 
                         upper = upper_bounds,
                         df_prediction = df.prediction_explanation)

print(fit.params$par)
```

#### Intervention only model

```{r}

initial_params <- c(temperature = 10)
lower_bounds <- c(temperature = 0)
upper_bounds <- c(temperature = 100)

fit.params_intervention_only <- optim(par = initial_params, 
                         fn = fun.fit_params_intervention_only,
                         method = "L-BFGS-B",
                         lower = lower_bounds, 
                         upper = upper_bounds,
                         df_prediction = df.prediction_explanation)

print(fit.params_intervention_only$par)
```

#### Truth only model

```{r}

initial_params <- c(temperature = 10)
lower_bounds <- c(temperature = 0)
upper_bounds <- c(temperature = 100)

fit.params_truth_only <- optim(par = initial_params, 
                         fn = fun.fit_params_truth_only,
                         method = "L-BFGS-B",
                         lower = lower_bounds, 
                         upper = upper_bounds,
                         df_prediction = df.prediction_explanation)

print(fit.params_truth_only$par)
```

### Predictions

#### Combined model 

```{r}
df.prediction_explanation = df.prediction_explanation %>%
  group_by(intervention, causal_structure, outcome) %>%
  mutate(prediction = fun.softmax(fit.params$par[1] * value + (1 - fit.params$par[1]) * truth, temp = fit.params$par[2])) %>%
  ungroup()

write.csv(df.prediction_explanation, file = "../../../data/model/explanation_predictions.csv", row.names = FALSE)
write.csv(df.prediction_intervention, file = "../../../data/model/intervention_predictions.csv", row.names = FALSE)
```

#### Intervention only model 

```{r}
df.prediction_explanation_intervention_only = df.prediction_explanation %>%
  group_by(intervention, causal_structure, outcome) %>%
  mutate(prediction = fun.softmax(value, temp = fit.params_intervention_only$par[1])) %>%
  ungroup()

write.csv(df.prediction_explanation_intervention_only, file = "../../../data/model/explanation_predictions_intervention_only.csv", row.names = FALSE)
```

#### Truth only model 

```{r}
df.prediction_explanation_truth_only = df.prediction_explanation %>%
  group_by(intervention, causal_structure, outcome) %>%
  mutate(prediction = fun.softmax(truth, temp = fit.params_truth_only$par[1])) %>%
  ungroup()

write.csv(df.prediction_explanation_truth_only, file = "../../../data/model/explanation_predictions_truth_only.csv", row.names = FALSE)
```



# Session info

```{r, echo=F}
sessionInfo()
```